\documentclass[english,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\usepackage[dvips]{graphicx}
\usepackage{tabularx}
\usepackage{array}
\usepackage{natbib}
\usepackage{delarray}
\usepackage{dcolumn}
\usepackage{float}
\usepackage{amsmath}
\usepackage{psfrag}
\usepackage{multirow,hhline}
\usepackage{amstext}
\usepackage{pstricks}
\usepackage{pst-plot}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage{cite}
\usepackage{rotating}
\usepackage{textcomp}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{pdflscape}
\usepackage{multirow}
\usepackage{bbding}
\usepackage{bbm}


\newcommand\doubleRule{\midrule\midrule}
\newcommand{\indep}{\perp \!\!\! \perp}
\newcommand{\notindep}{\not\indep}

\newcommand{\pr}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\re}[1]{\smallskip\textsc{Respuesta} \begin{it}\\ #1 \end{it} \bigskip}
\textwidth=480pt \textheight=610pt
\oddsidemargin=0in\topmargin=-.2in
\def\max{mathop{\mbox{\normalfont m\'ax}}\limits}
\def\min{mathop{\mbox{\normalfont m\'{i}n}}\limits}

\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}


\usepackage{multirow,hhline}



\begin{document}



\title{\sc Empirical Industrial Organization\\Fall 2022 - Take Home Exam 2}
\author{Instructor: Cristi\'an S\'anchez. \\ Student: Jes\'us L\'opez P\'erez 200863}
\date{Due date: December 13, 2022}

\maketitle


\bigskip
\bigskip


%%%%%%%%%%%%%
\noindent In this exam we investigate a static continuous game with complete information in the supply side. We consider two scenarios: one where the firms have constant marginal costs, and another one where firms have increasing marginal costs. We study the solution and estimation of this game.


\section{Constant Marginal Costs} 
Suppose that we have already estimated a demand model, thus we have estimates for the demand function of the firm, $q_j(p)$. Suppose, too, that the profit function of the firm is,
\[ \Pi_j(p) = p_jq_j(p) - c_j\left(q_j(p)\right) , \]
where $c_j\left(q_j(p)\right) = c_0 + c_1q_j(p) + Z_j\gamma + \nu_j$, in which $c_0$, $c_1$, and $\gamma$ are unknown (to the econometrician) parameters to be estimated, $Z_j$ is a vector of cost shifters, and $\nu_j$ is a mean zero unobserved (to the econometrician) determinant of costs.
\begin{enumerate}
	\item Explicitly write down the expression for the marginal cost, 
	$c_j^\prime(q_j)$. \\
	\textit{R = }
	\[
mc = 	c_j^\prime (q_j) = \frac{\partial c_j\left(q_j(p)\right)}{ \partial p_j} = c_1 
	\]
	
	\item Using your expression for $c_j^\prime(q_j)$, write down the first order condition for firm $j$ that chooses $p_j$ to maximize profits.\\
\textit{R =} \\
Firms set each of their prices to maximize 

\begin{align*}
\sum_{j \in J} p_j q_j -c_j\left(q_j(p)\right) &= p_j q_j - c_0 - c_1q_j(p) - Z_j\gamma - \nu_j \\
& =  \sum_{j \in J} \left( p_j - c_1 \right) q_j(p) - c_0 -Z_j \gamma - \nu_j\\
& = \sum_{j \in J} \left( p_j - c_1 \right) M s_j - c_0 -Z_j \gamma - \nu_j
\end{align*}

where $M$ is the total market and $s_j$ is the firm share. Thus implies that for $j = 1, \cdots ,J$ 
\begin{align*}
s_j + \sum_{k \in J_f} (p_k - mc_k) M \frac{\partial s_k}{\partial p_j} &= 0 \\
s_j + \sum_{k \in J_f} (p_k - c_1) M \frac{\partial s_k}{\partial p_j} &= 0 \\
\end{align*}
	
	\item Using the estimated demand function, $\hat q_j(p)$, the expression for the firm's first order condition, and assuming no endogeneity concerns, how would you estimate the cost function parameters, $c_0$, $c_1$, and $\gamma$?\\
\textit{R = }\\
Note that we can rewrite the FOC in matrix form

\[
\Omega(p)(p-mc) + s(p) = 0
\]

Where $\Omega(p)$ is the matrix of $\frac{\partial s_k}{\partial p_j} $ effects, where values are nonzero for elements of a row that are owned by the same firm as the row good. We solve for $mc = p - \Omega^{-1}(p) s(p)$

Then, we follow the next steps
\begin{enumerate}
\item Estimate the random coefficients model (RC).
\item Calculate $\Omega(p)$ at observed prices and estimated coefficients.
\item Back out $mc$
\[
\hat{mc} = p - \Omega^{-1}(p, \hat{\theta}) s
\]
\end{enumerate}	
	
	\item Notice that you would need to adjust your standard errors, because you are using an estimate of $q_j(p)$, which induces measurement error in your estimation. How would you adjust your standard errors in this context? \\
	
\textit{R=} I might use the delta method or a bootstrap to calculate standard errors. Note that bootstrap is preferred to avoid inverse matrix calculations.
	
	\item Now, suppose that you have your supply side estimates at hand, i.e. $\hat c_0$, $\hat c_1$, $\hat\gamma$. Suppose, too, that there are only two firms in the market, $j=1,2$, and that you want to simulate new equilibrium prices, $p^*$, that correspond to new values of the vector of cost shifters, $Z^\prime$. Write down the system of (two) equations that you would need to solve to obtain the new equilibrium prices. \\
\textit{R = }
\[
p^* = \hat{mc} + \Omega^{post,-1} (p^*, \hat{\theta}) s(p^*, \hat{\theta})
\]	
	
\end{enumerate}



\section{Increasing Marginal Costs}
Now, suppose that the cost function is quadratic, and of the form
\[ c_j\left(q_j(p)\right) = c_0 + c_1q_j(p) + c_2q_j(p)^2 + Z_j\gamma + \nu_j , \]
where $c_0$, $c_1$, $c_2$, and $\gamma$ are unknown (to the econometrician) parameters to be estimated.
\begin{enumerate}
	\item Explicitly write down the expression for the marginal cost, $c_j^\prime(q_j)$.\\
\textit{R = }\\
	\[
mc = 	c_j^\prime (q_j) = \frac{\partial c_j\left(q_j(p)\right)}{ \partial p_j} = c_1+ 2 c_2 q_j(p)
	\]
	
	\item Using your expression for $c_j^\prime(q_j)$, write down the first order condition for firm $j$ that chooses $p_j$ to maximize profits.\\
\textit{R = }	\\
Thus implies that for $j = 1, \cdots ,J$ 
\begin{align*}
	s_j + \sum_{k \in J_f} (p_k - mc_k) M \frac{\partial s_k}{\partial p_j} = 0 \\
	s_j + \sum_{k \in J_f} (p_k - c_1+ 2 c_2 q_j(p)) M \frac{\partial s_k}{\partial p_j} = 0
\end{align*}

	
	\item Using the estimated demand function, $\hat q_j(p)$, the expression for the firm's first order condition, and assuming no endogeneity concerns, how would you estimate the cost function parameters, $c_0$, $c_1$, and $\gamma$?
\end{enumerate}



%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document} 
